---
title: "UNSUPERVISEDLEARNING WITH R"
author: "ROTICH HARON"
date: "1/28/2022"
output: html_document
---
# Defining the question:
```{r}
#to understand the customer’s behavior  and characteristics of  customer groups"
```
#Metrics of success
```{r}
#idenify charactersistcs of various customer groups and clustering them
```
# Context
```{r}
#Kira Plastinina  is a Russian brand that is sold through a defunct chain of #retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and #Armenia,The brand’s Sales and Marketing team would like to understand their #customer’s behavior from data that they have collected over the past year.
```
#Experimental Design
```{r}
#create a Markdown file  to perform data exploration , cleaning, univariate and #Bivariate analysis and clustering
```
---

```{r}
#install.packages("tidyverse")
library(tidyverse)
```

```{r}
# loading dataset
df<- read.csv("C:/Users/hp/Downloads/online_shoppers_intention.csv")
# previewing data set
head(df)
```
```{r}
# checking shape of data
dim(df)
# our data set has 12330 rows and 18 columns 
```
```{r}
# finding summary of our data
summary(df)
```
```{r}
# checking for  data types
str(df)
```
```{r}
# checking for missing values
sum(is.na(df))
# there are 112 missing values

```
```{r}
# dealing with missing values
df1<-na.omit(df)

```
```{r}
sum(is.na(df1))
```

```{r}
# checking for duplicates
sum(duplicated(df1))
# there are 117 duplicates
```
```{r}
library(dplyr) 
# dealing with duplicated rows
df1 %>% distinct()
```
```{r}
# outlier detection
# Simple boxplot  for perating syatems
boxplot(df1$OperatingSystems)
# there a number of outliers in OperatingSystems
```
```{r}
# outlier detection in special day variable
boxplot(df1$SpecialDay)
# there are outlier present in SpecialDay
```
```{r}

# outlier detetion in page values
boxplot(df1$PageValues)
#there are many outliers in this variable
```
```{r}

# outlier detection for ExitRates
boxplot(df1$ExitRates)
#this variable has  a large number of outliers

```
```{r}
# Simple boxplot of Bouncerates
boxplot(df1$BounceRates)
#
```

```{r}

# Simple boxplot of product related_ Duration
boxplot(df1$ProductRelated_Duration)
#
```
```{r}

# Simple boxplot of ProductRelated

boxplot(df1$ProductRelated)
# there a number of outliers in this variable
```
```{r}

# outlier detection forInformational_Duration
boxplot(df1$Informational_Duration)
# Informational_Duration has  outliers
```
```{r}

# 'Administrative_Duration outlier detection
boxplot(df1$Administrative_Duration)

# 'Administrative_Duration has outliers
```
###
# UNIVARIATE ANALYSIS
## measures of central tendencies

```{r}

# operating systems
#find mean
mean(df1$OperatingSystems)
```
```{r}
# median
median(df1$OperatingSystems)
```
```{r}
##find range
max(df1$OperatingSystems) - min(df1$OperatingSystems)
```
```{r}
##find interquartile range (spread of middle 50% of values)
IQR(df1$OperatingSystems)
```
```{r}
##find standard deviation
sd(df1$OperatingSystems)
```

# specialDay
```{r}
#find mean
mean(df1$SpecialDay)

```
```{r}
median(df1$SpecialDay)
```
```{r}
##find range
max(df1$SpecialDay) - min(df1$SpecialDay)
```
```{r}
#find standard deviation
sd(df1$OperatingSystems)
```
## page values
```{r}
# finding mean
mean(df1$PageValues)
```
```{r}
# finding  median
median(df1$PageValues)
```
```{r}
# finding  range
max(df1$PageValues)- min(df1$PageValues)
```
```{r}
# std
sd(df1$PageValues)
```
#
#Exit rates
```{r}
# mean
mean(df1$ExitRates)
```

```{r}
# median
median(df1$ExitRates)
```
```{r}
# range
max(df1$ExitRates)- min(df1$ExitRates)
```
```{r}
# std
sd(df1$ExitRates)
```

## Bounce rates
```{r}
# mean
mean(df1$BounceRates)
```
```{r}
# median
median(df1$BounceRates)
```
```{r}
# range
max(df1$BounceRates)-min(df1$BounceRates)
```
```{r}
# std
sd(df1$BounceRates)
```

## ProductRelation Duration
```{r}
# mean
mean(df1$ProductRelated_Duration)
```
```{r}
# median
median(df1$ProductRelated_Duration)
```
```{r}
# range
max(df1$ProductRelated_Duration)- min(df1$ProductRelated_Duration)
```
```{r}
#std
sd(df1$ProductRelated_Duration)
```
##
#
# Product related
```{r}
# mean
mean(df1$ProductRelated)
```
```{r}
median(df1$ProductRelated)
```
```{r}
# range
max(df1$ProductRelated)- min(df1$ProductRelated)
```
```{r}
# std
sd(df1$ProductRelated)
```
##
# Informational_Duration
```{r}
#mean
mean(df1$Informational_Duration)
```
```{r}
# median
median(df1$Informational_Duration)
```
```{r}
# range
max(df1$Informational_Duration)- min(df1$Informational_Duration)
```
```{r}
# std
sd(df1$Informational_Duration)
```

# Administrative Duration
```{r}
#mean
mean(df1$Administrative_Duration)
```
```{r}
#median
median(df1$Administrative_Duration)
```
```{r}
# range
max(df1$Administrative_Duration)- min(df1$Administrative_Duration)
```
```{r}
#std
sd(df1$Administrative_Duration)
```
## Administrative
```{r}
# mean
mean(df1$Administrative)
```
```{r}
#median
median(df1$Administrative)
```
```{r}
# range

max(df1$Administrative)- min(df1$Administrative)
```
```{r}
#std
sd(df1$Administrative)
```
#Histogram
```{r}
#produce histogram
A <- df1$Administrative
hist(A)
# There is high visits on  administration  at category 0
```
```{r}
#produce histogram
BounceRates <- df1$BounceRates
hist(BounceRates)

```

```{r}
# produce histogram
PageValues <- df1$PageValues
hist(PageValues)
```
```{r}
productrelated <- df1$ProductRelated
hist(productrelated)
# most visits are not product related
```
# frequency table
```{r}
freq <- df1$Informational
table(freq)
# majority of the visit  have informational level of 0
```
```{r}
A <- df1$Administrative
table(A)
# most visits are non administrative
```
```{r}
X <- df1$SpecialDay
table(X)
# most visits are done away from special days 
```
```{r}
Y <- df1$Browser
table(Y)
# Browser 1 is mostly used to make visits
```
```{r}
region <- df1$Region
table(region)
# region  1 has the highest visits
```
```{r}
traffic <- df1$TrafficType
hist(traffic)
# the traffict type that has most visits are between 1-4
```
```{r}
os <- df1$OperatingSystems
hist(os)
# the visitors using  the operating system 2 are te majority
```
```{r}
Vtype<- df1$VisitorType
table(Vtype)
# the returning vistors are majority of those who make visits
```
```{r}
weekend<- df1$Weekend
table(weekend)
# most vists are made on weekdays than weekends
```

# BIVARIATE ANALYSIS
### scatter plots
```{r}
#create scatterplot of SpecialDaysvs. PageDAYS
plot(df1$SpecialDay, df1$PageValues, pch=16, col='steelblue',
     main='SpecialDay vs. PageValues',
     xlab='SpecialDay', ylab='PageValues')
# the visits  far away from special days there are  more or high page values

```

```{r}
#create scatterplot of Revenue vs. Administrative
plot(df1$Revenue, df1$Administrative, pch=16, col='steelblue',
     main='revenue vs. Administrative',
     xlab='revenue', ylab='Administrative')
#there isn't any relationshp between revenue and  administrative
```
```{r}
#create scatterplot ofrevenues vs. weekend
plot(df1$Administrative, df1$Administrative_Duration, pch=16, col='steelblue',
     main='Administrative vs. Administrative duration',
     xlab='Adminiatrative', ylab='AdministrativeDuration')
```

```{r}
#create scatterplot of ProductRelated_Duration vs.ProductRelated
plot(df1$ProductRelated, df1$ProductRelated_Duration, pch=16, col='steelblue',
     main='ProductRelated_Duration vs.ProductRelated',
     xlab='ProductRelated', ylab='ProductRelated_Duration')

```
```{r}
#create scatterplot of ProductRelated_Duration vs.ProductRelated
plot(df1$ProductRelated, df1$Administrative, pch=16, col='steelblue',
     main='Administrative vs.ProductRelated',
     xlab='product related', ylab='Admninistrative')
```
```{r}
#create scatterplot of ProductRelated_Duration vs.ProductRelated
plot(df1$ProductRelated, df1$BounceRates, pch=16, col='steelblue',
     main='BounceRates vs.ProductRelated',
     xlab='ProductRelated', ylab='Bounce rates')
# there is a positive realationship between bounce rate and  product related variable
```
```{r}
#create scatterplot of ProductRelated_Duration vs.BounceRATES
plot(df1$BounceRates, df1$ProductRelated_Duration, pch=16, col='steelblue',
     main='ProductRelated_Duration vs.Bounce Rates',
     xlab='BounceRate', ylab='ProductRelated_Duration')
# there is a positive relationship between  product related and bounce rates
```
```{r}
#create scatterplot of ProductRelated_Duration vs.ProductRelated
plot(df1$BounceRates, df1$ExitRates, pch=16, col='steelblue',
     main='BounceRates vs.Exitrates',
     xlab='BounceRates', ylab='ProductRelated_Duration')

```
```{r}
cor(df1[, unlist(lapply(df1, is.numeric))])

```
##encoding categorical
```{r}
library(caret)

dmy <- dummyVars(" ~ .", data = df1, fullRank = T)
df1_new <- data.frame(predict(dmy, newdata = df1))

```

## Mulvariate analysis : PCA
```{r}
pca.fit <- prcomp(x = df1_new, scale. = TRUE)

```
```{r}
pca.summary <- summary(pca.fit)
ls(pca.summary)
```
```{r}
# to check on the important factors
pca.summary$importance
#The second row, Proportion of Variance, shows how much variation in the data is described by each component; notice that the first component, PC1, explains the most variance, 0.132170 of the total variance, the second component explains the second most variance is 0.0743, and so on, with each successive component explaining a lower proportion of the total variance
```
## IMplementing the solution
### kmeans clustering
```{r}
# removing label column
d<- df1_new[, c(1, 2, 3, 4,5,6,7,8,9,10,11,12,13,14,15,16,17)]
#d.class<- df1_new[, "Revenue"]
head(d)

```
```{r}
# normalizing data
#define Min-Max normalization function
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }

#apply Min-Max normalization to first four columns in iris dataset
d_norm <- as.data.frame(lapply(d[1:17], min_max_norm))
head(d_norm)
```

```{r}
# implementing Kmeans where k is 4
result<- kmeans(d_norm,4) 
result$size 

```
```{r}
# to visualize results
par(mfrow = c(1,2), mar = c(5,4,2,2))
```
```{r}
plot(d_norm[,5:6], col = result$cluster) 
```

# DBSCAN
```{r}
# to find optimal eps using K-NN distplot
dbscan::kNNdistplot(d_norm,k=4)
abline(h=0.15,lty = 2,col=rainbow(1),main="eps optimal value")
```
```{r}
## applying the DBSCAN 
library("dbscan")
db<-dbscan(d_norm,eps=0.2,MinPts = 4)
db
```
```{r}
hullplot(d_norm,db$cluster)
```
### HIERACHICAL CLUSTEERING
```{r}
library(cluster)

```
```{r}
df1_new <- scale(d_norm)
```
```{r}
d <- dist(df1_new, method = "euclidean")

```
```{r}
res.hc <- hclust(d, method = "ward.D2" )
res.hc
```
```{r}
plot(res.hc, cex = 0.6, hang =-1)
```

## conclusion and recomendation
```{r}
##1.most visits are non administrative thus more efforts should be made on nod administrative pages.
##2.the returning visitors are majority of those who make hence  more efforts should be made to satisfy   existing  customers since they are majority, while also investing more on the new customers.
#3.majority of the visit  have informational level of 0
#4.there is a positive relationship between bounce rate and  product related variable, more strategies should be implemented on product related sites.
#5.most visits are done away from special days , with this  the team should come up with more strategies to  target marketing on days way from  special days.
#6. more marketing efforts should be made on weekdays  because most visits are made on such day
#7. For clustering the DBSCAN prove to be more good  for clustering
```
---